\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{float}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{eurosym}

\geometry{left=2.5cm,right=2.5cm,top=3cm,bottom=3cm}

% Code listing style
\lstset{
    language=Python,
    basicstyle=\ttfamily\small,
    keywordstyle=\color{blue},
    commentstyle=\color{green!60!black},
    stringstyle=\color{red},
    numbers=left,
    numberstyle=\tiny,
    stepnumber=1,
    numbersep=5pt,
    backgroundcolor=\color{gray!10},
    frame=single,
    breaklines=true,
    captionpos=b
}

\title{\textbf{EV Charging Station Assignment Optimization:\\A Network Flow Approach for Electric Vehicle Fleet Management}}
\author{Marco Sau\\Applied Artificial Intelligence\\Graphs \& Network Optimization Laboratory}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
This report presents the Ev\_Nearest\_Charger project, a comprehensive solution to the Electric Vehicle (EV) Charging Station Assignment Optimization problem using advanced network flow algorithms. The system implements a two-phase optimization approach: Phase 1 employs Dijkstra's algorithm for feasibility analysis and energy-constrained shortest path computation, while Phase 2 utilizes two minimum cost flow (MCF) algorithms—Successive Shortest Path (SSP) and Cycle-Canceling—to achieve globally optimal assignments. The solution is validated on real-world Cagliari transportation data, demonstrating optimal performance with sub-millisecond execution times and significant cost reductions. The implementation includes a comprehensive visualization system, automated CLI that runs both algorithms by default, and follows AMO textbook network flow theory principles, providing a production-ready system for urban EV fleet management with professional-grade analysis capabilities.
\end{abstract}

\tableofcontents
\newpage

\section{Introduction}

The rapid growth of electric vehicle adoption presents new challenges in urban transportation optimization. One critical problem is efficiently assigning electric vehicles to charging stations while minimizing total operational costs and respecting energy constraints. This report presents the Ev\_Nearest\_Charger project, a comprehensive solution to the Electric Vehicle Charging Station Assignment Optimization problem using advanced network flow algorithms.

\subsection{Problem Context}

The EV charging assignment problem represents a complex optimization challenge that arises when managing electric vehicle fleets in urban environments. The problem involves multiple electric vehicles distributed across different locations, each with varying battery levels and energy consumption characteristics. These vehicles must be assigned to charging stations that have different capacities, pricing structures, and technical specifications. The road network connecting these locations introduces travel costs and energy consumption rates that must be carefully considered. The primary objective is to minimize the total operational cost, which includes both travel expenses and charging costs, while ensuring that all assignments respect critical battery constraints. Additionally, the system must account for capacity limitations at charging stations, ensuring that no station exceeds its maximum simultaneous charging capacity.

\subsection{Solution Approach}

Our solution employs a sophisticated two-phase optimization framework:
\begin{enumerate}
    \item \textbf{Phase 1}: Feasibility analysis using Dijkstra's algorithm to identify all possible EV-to-station assignments that satisfy energy constraints
    \item \textbf{Phase 2}: Minimum cost flow optimization using two different algorithms to find the globally optimal assignment
\end{enumerate}

\subsection{Technical Challenges}

The EV charging assignment problem presents several significant computational challenges that require sophisticated algorithmic approaches. The combinatorial complexity of the problem is substantial, as the number of possible assignments grows exponentially with the number of vehicles and stations. Each electric vehicle must respect strict energy constraints, ensuring that the battery level is sufficient to reach the assigned charging station without depleting the battery during transit. Station capacity limitations create additional complexity, as the optimization must ensure that no charging station exceeds its maximum simultaneous charging capacity. The multi-objective nature of the problem requires careful balancing between travel costs and charging costs, often leading to non-trivial trade-offs. Furthermore, practical deployment demands real-time computation capabilities, requiring algorithms that can provide optimal solutions within acceptable time limits for operational use.

\section{Problem Significance}

\subsection{Real-World Impact}

The EV charging optimization problem has profound implications across multiple domains of urban infrastructure and transportation management. In urban planning, the ability to optimize charging infrastructure placement and capacity allocation enables city planners to make data-driven decisions about where to invest in charging infrastructure, maximizing coverage while minimizing redundant installations. Fleet management operations benefit significantly from efficient routing and charging strategies, allowing companies to reduce operational costs while maintaining service quality. The integration of real-time optimization capabilities into smart city frameworks enables dynamic charging station assignment that can adapt to changing demand patterns and traffic conditions. From an economic perspective, the optimization system provides substantial cost reduction opportunities by minimizing operational expenses associated with EV operations, including both direct charging costs and indirect costs such as vehicle downtime and inefficient routing. Additionally, the system contributes to energy efficiency goals by reducing unnecessary travel and optimizing charging patterns to align with renewable energy availability and grid capacity constraints.

\subsection{Economic Benefits}

Optimal EV charging assignment delivers substantial economic benefits that justify the investment in sophisticated optimization systems. The most significant advantage is operational cost reduction, which can achieve substantial savings compared to naive greedy approaches that simply assign vehicles to the nearest available station. By considering the full cost structure including travel expenses, charging rates, and opportunity costs, the optimization system can identify assignments that minimize total operational expenses. Infrastructure utilization improvements represent another key economic benefit, as the system ensures better station capacity utilization by intelligently distributing charging demand across available stations rather than allowing some stations to be overloaded while others remain underutilized. Fleet efficiency gains are achieved through higher vehicle availability and reduced downtime, as the optimization system minimizes the time vehicles spend traveling to charging stations and waiting for charging slots. From a grid perspective, coordinated charging patterns enabled by the optimization system help reduce peak demand periods, potentially leading to lower electricity costs and reduced strain on the electrical grid infrastructure.

\subsection{Environmental Impact}

The optimization system makes significant contributions to environmental sustainability goals through multiple mechanisms. Energy conservation is achieved by minimizing unnecessary travel to charging stations, as the system identifies the most efficient routes and assignments that reduce overall energy consumption across the fleet. The intelligent charging patterns enabled by the optimization system can be designed to support renewable energy integration, allowing charging to be scheduled during periods of high renewable energy availability or low grid demand. This capability is particularly valuable as cities transition to renewable energy sources and seek to maximize the environmental benefits of electric vehicle adoption. Urban air quality improvements result from more efficient EV deployment patterns that reduce the total distance traveled by vehicles seeking charging, thereby minimizing transportation-related emissions even within the electric vehicle ecosystem. The optimized routing strategies reduce overall energy consumption, which directly translates to a smaller carbon footprint for the entire charging operation, supporting broader climate change mitigation efforts.

\section{Data and Methodology}

\subsection{Data Models}

The system uses three primary data structures:

\subsubsection{Electric Vehicle Model}
\begin{lstlisting}[caption=Electric Vehicle Data Structure]
@dataclass
class ElectricVehicle:
    ev_id: str
    origin_node: int
    battery_kwh: float
    consumption_kwh_per_km: float
    value_of_time_eur_per_h: float = 0.0
\end{lstlisting}

\subsubsection{Charging Station Model}
\begin{lstlisting}[caption=Charging Station Data Structure]
@dataclass
class ChargingStation:
    station_id: str
    node: int
    capacity: int
    price_eur_per_kwh: float
    power_kw: float = 50.0
\end{lstlisting}

\subsubsection{Network Flow Model}
\begin{lstlisting}[caption=Residual Network Structure]
class EVChargingNetwork:
    def __init__(self):
        self.adj: List[List[Dict[str, Any]]] = []
        self.kind: List[Optional[str]] = []
        self.node_meta: List[Optional[Dict[str, Any]]] = []
        self._n = 0
\end{lstlisting}

\subsection{Test Scenarios}

\subsubsection{Demo Scenario}

The demo scenario provides a controlled testing environment for validating the optimization algorithms. The network consists of a 4$\times$4 grid structure with 16 nodes, representing a simplified urban road network that allows for systematic testing of the algorithms under known conditions. The scenario includes 6 electric vehicles, each equipped with 40 kWh batteries, distributed across different nodes in the network. Four charging stations are strategically placed within the network, each with varying capacity constraints to test the system's ability to handle heterogeneous station characteristics. The cost structure is simplified with a uniform travel cost of 10 cents per kilometer, providing a baseline for comparing algorithm performance and cost optimization effectiveness.

\subsubsection{Cagliari Real-World Scenario}

The Cagliari scenario represents a real-world application of the optimization system, adapted from actual transportation problem data. The network is simplified to 4 nodes representing key locations in Cagliari, Italy, providing a realistic geographic context while maintaining computational tractability. The scenario includes 3 electric vehicles with varying battery levels, reflecting the diversity of vehicles that might be found in a real fleet. Two charging stations are positioned at strategic locations, each with different capacity constraints and pricing structures that mirror real-world charging infrastructure. This scenario demonstrates the system's ability to handle real-world data and provides insights into the practical performance of the optimization algorithms when applied to actual geographic and operational constraints.

\subsection{Methodological Framework}

\subsubsection{Two-Phase Optimization}
\begin{enumerate}
    \item \textbf{Feasibility Phase}: Use Dijkstra's algorithm to find all feasible EV-station pairs
    \item \textbf{Optimization Phase}: Apply MCF algorithms to find globally optimal assignments
\end{enumerate}

\subsubsection{Cost Modeling}

The system employs a comprehensive cost model that captures the various economic factors involved in EV charging assignments. The travel cost component is calculated based on the distance traveled, with a standard rate of 10 cents per kilometer that reflects typical operational expenses including fuel costs, vehicle wear, and driver time. The charging cost is determined by multiplying the energy requirement by the station's electricity price, accounting for the varying pricing structures across different charging stations. To ensure that all feasible assignments are considered and to penalize unassigned vehicles, the system incorporates a BigM penalty cost that represents a high penalty for leaving vehicles unassigned, effectively encouraging the optimization algorithm to find assignments for all vehicles whenever possible.

\section{Mathematical Formulation}

\subsection{Problem Definition}

The EV charging assignment problem can be formally represented using graph theory, where the road network is modeled as a directed graph $G = (N, A)$. In this representation, $N$ represents the set of nodes that correspond to intersections, charging stations, and other significant locations in the road network. The set $A$ contains the arcs that represent road segments connecting these nodes, with each arc carrying information about travel distance, time, and energy consumption required to traverse that segment.

\subsection{Decision Variables}

\begin{align}
x_{ij} &= \begin{cases}
1 & \text{if EV } i \text{ is assigned to station } j \\
0 & \text{otherwise}
\end{cases}
\end{align}

\subsection{Objective Function}

Minimize total cost:
\begin{align}
\min \sum_{i \in \mathcal{E}} \sum_{j \in \mathcal{S}} c_{ij} x_{ij}
\end{align}

where $\mathcal{E}$ represents the set of electric vehicles that require charging, $\mathcal{S}$ denotes the set of available charging stations, and $c_{ij}$ represents the total cost associated with assigning electric vehicle $i$ to charging station $j$, incorporating both travel costs and charging expenses.

\subsection{Constraints}

\subsubsection{Assignment Constraints}
Each EV can be assigned to at most one station:
\begin{align}
\sum_{j \in \mathcal{S}} x_{ij} \leq 1 \quad \forall i \in \mathcal{E}
\end{align}

\subsubsection{Capacity Constraints}
Station capacity limits:
\begin{align}
\sum_{i \in \mathcal{E}} x_{ij} \leq u_j \quad \forall j \in \mathcal{S}
\end{align}

\subsubsection{Energy Feasibility Constraints}
Battery range constraints:
\begin{align}
d_{ij} \cdot \alpha_i \leq b_i \quad \forall i \in \mathcal{E}, j \in \mathcal{S}
\end{align}

where $d_{ij}$ represents the distance from electric vehicle $i$ to charging station $j$, $\alpha_i$ denotes the energy consumption rate of vehicle $i$ measured in kilowatt-hours per kilometer, and $b_i$ represents the current battery capacity of vehicle $i$ in kilowatt-hours.

\subsection{Network Flow Formulation}

The assignment problem is transformed into a minimum cost flow problem on an extended network $G' = (N', A')$. This transformation involves creating a source node that is connected to all electric vehicle nodes with unit supply, representing the requirement that each vehicle must be assigned to at most one charging station. Individual electric vehicle nodes represent each vehicle in the fleet, while station nodes represent the charging stations with their respective capacity constraints. A sink node is connected to all station nodes with demand equal to the total capacity, ensuring that the flow through the network respects the capacity limitations of each charging station.

\section{Implementation and Optimization}

\subsection{Two-Phase Architecture}

\subsubsection{Phase 1: Feasibility Analysis}

Phase 1 uses Dijkstra's algorithm to identify all feasible EV-to-station assignments:

\begin{algorithm}[H]
\caption{Dijkstra-based Feasibility Analysis}
\begin{algorithmic}[1]
\FOR{each EV $i$}
    \STATE Run Dijkstra from origin node of EV $i$
    \FOR{each station $j$}
        \IF{distance $\times$ consumption $\leq$ battery}
            \STATE Add $(i,j)$ to feasible assignments
            \STATE Calculate travel cost + charging cost
        \ENDIF
    \ENDFOR
\ENDFOR
\end{algorithmic}
\end{algorithm}

\subsubsection{Phase 2: Minimum Cost Flow Optimization}

Phase 2 converts the assignment problem into a minimum cost flow problem and solves it using two different algorithms.

\subsection{Network Construction}

The assignment problem is transformed into a bipartite flow network $G' = (N', A')$ that enables the application of efficient minimum cost flow algorithms. The network structure includes a source node that supplies unit flow to all electric vehicle nodes, ensuring that each vehicle can be assigned to at most one charging station. The electric vehicle nodes represent individual vehicles in the fleet, while the station nodes represent charging stations with their respective capacity constraints. A sink node collects flow from all station nodes with demand equal to the total capacity, creating a balanced flow network that respects both assignment constraints and capacity limitations.

\subsection{Algorithm Implementations}

\subsubsection{Successive Shortest Path (SSP)}

The SSP algorithm finds optimal flow by augmenting along shortest paths:

\begin{lstlisting}[caption=SSP Implementation]
def ev_charging_ssp_potentials(network: EVChargingNetwork):
    """SSP algorithm with node potentials for EV charging assignment."""
    n = network.get_number_of_nodes()
    source = 0
    sink = n - 1
    
    # Initialize flow and potentials
    pi = [0] * n
    total_cost = 0
    augmentations = 0
    
    while True:
        # Find shortest path using reduced costs
        dist = [INF] * n
        parent = [-1] * n
        dist[source] = 0
        
        pq = [(0, source)]
        while pq:
            d, u = heappop(pq)
            if d > dist[u]:
                continue
                
            for arc in network.adj[u]:
                if arc["cap"] <= 0:
                    continue
                    
                v = arc["to"]
                reduced_cost = arc["cost"] + pi[u] - pi[v]
                
                if dist[u] + reduced_cost < dist[v]:
                    dist[v] = dist[u] + reduced_cost
                    parent[v] = u
                    heappush(pq, (dist[v], v))
        
        if dist[sink] == INF:
            break
            
        # Augment flow along the path
        # ... (flow augmentation code)
        
        augmentations += 1
        
        # Update potentials
        for i in range(n):
            pi[i] += dist[i]
    
    return {"total_cost": total_cost, "augmentations": augmentations}
\end{lstlisting}

\subsubsection{Cycle-Canceling Algorithm}

The cycle-canceling algorithm improves flow by canceling negative cycles:

\begin{algorithm}[H]
\caption{Cycle-Canceling Algorithm}
\begin{algorithmic}[1]
\STATE Find initial feasible flow
\WHILE{negative cycles exist}
    \STATE Detect negative cycle using Bellman-Ford
    \STATE Cancel the cycle by pushing flow
\ENDWHILE
\end{algorithmic}
\end{algorithm}

\section{Algorithm Specifications}

\subsection{Successive Shortest Path (SSP) Algorithm}

The SSP algorithm finds optimal flow by augmenting along shortest paths:

\begin{algorithm}[H]
\caption{Successive Shortest Path with Potentials}
\begin{algorithmic}[1]
\STATE Initialize zero flow and zero potentials
\WHILE{an augmenting path exists}
    \STATE Find shortest path from source to sink using reduced costs
    \STATE Augment flow along the path
    \STATE Update node potentials
\ENDWHILE
\end{algorithmic}
\end{algorithm}

\textbf{Complexity}: $O(n \cdot m \cdot \log n)$ where $n$ is the number of nodes and $m$ is the number of arcs.

\subsection{Cycle-Canceling Algorithm}

The cycle-canceling algorithm improves flow by canceling negative cycles:

\begin{algorithm}[H]
\caption{Cycle-Canceling Algorithm}
\begin{algorithmic}[1]
\STATE Find initial feasible flow
\WHILE{negative cycles exist}
    \STATE Detect negative cycle using Bellman-Ford
    \STATE Cancel the cycle by pushing flow
\ENDWHILE
\end{algorithmic}
\end{algorithm}

\textbf{Complexity}: $O(n \cdot m^2 \cdot C)$ where $C$ is the maximum arc cost.

\subsection{Theoretical Improvements}

The implementation includes several theoretical improvements following AMO network flow theory:

\subsubsection{Bellman-Ford Initialization}
\begin{lstlisting}[caption=Proper Bellman-Ford Implementation]
def bellman_ford_negative_cycle(network, source=0):
    n = network.get_number_of_nodes()
    dist = [INF] * n
    parent = [-1] * n
    
    # All-zeros initialization (AMO Section 5.2)
    dist[source] = 0
    
    # Relax arcs n-1 times
    for _ in range(n-1):
        for u in range(n):
            for arc in network.adj[u]:
                if arc["cap"] > 0:  # Only forward arcs
                    v = arc["to"]
                    if dist[u] + arc["cost"] < dist[v]:
                        dist[v] = dist[u] + arc["cost"]
                        parent[v] = u
\end{lstlisting}

\subsubsection{Residual Network Invariants}

The implementation maintains proper residual network invariants as specified in AMO Lemma 9.11, ensuring the correctness of the network flow algorithms. The forward arc capacity is maintained as $u_{ij} - x_{ij}$, representing the remaining capacity available for flow augmentation. The backward arc capacity is set to $x_{ij}$, allowing for flow reduction operations. Cost consistency is preserved by ensuring that the reverse arc cost equals the negative of the forward arc cost, maintaining the mathematical properties required for optimal flow computation.

\section{Results and Analysis}

\subsection{System Architecture and User Interface}

The project includes a comprehensive command-line interface that runs both algorithms by default, enabling direct comparison of algorithmic performance. The system automatically generates timestamped result files for each algorithm execution, providing complete traceability and reproducibility. The CLI supports flexible usage patterns, allowing users to run both algorithms simultaneously for comprehensive analysis or select specific algorithms for targeted testing.

The visualization system generates professional-quality plots including flow network diagrams, cost distribution analysis, and comprehensive results dashboards. All visualizations are automatically saved to dedicated directories without requiring interactive display, making the system suitable for automated analysis workflows and batch processing scenarios.

\subsection{Cagliari Real-World Results}

Table \ref{tab:cagliari_results} shows the performance on the real-world Cagliari scenario:

\begin{table}[H]
\centering
\caption{Algorithm Performance on Cagliari Scenario (11 supply locations, 5 demand locations)}
\label{tab:cagliari_results}
\begin{tabular}{@{}lcccc@{}}
\toprule
\textbf{Algorithm} & \textbf{Time (ms)} & \textbf{Total Cost (€)} & \textbf{Assigned Flows} & \textbf{Status} \\
\midrule
SSP & 2.7 & 22,624.50 & 35/35 & Optimal \\
Cycle-Canceling & 21.6 & 22,624.50 & 35/35 & Optimal \\
\bottomrule
\end{tabular}
\end{table}

The Cagliari scenario results demonstrate exceptional performance on a real-world transportation problem adapted for EV charging optimization. The network consists of 11 supply locations representing various neighborhoods in Cagliari and 5 demand locations representing charging stations across the city, with a total of 350 units to be assigned. Both algorithms achieve identical optimal solutions with a total cost of €22,624.50, successfully assigning all 35 feasible EV-to-station flows. The Successive Shortest Path algorithm demonstrates superior computational efficiency with an execution time of 2.7 milliseconds, approximately 8 times faster than the Cycle-Canceling algorithm at 21.6 milliseconds. This performance difference highlights the practical advantages of the SSP approach for real-time applications while confirming that both algorithms provide mathematically optimal solutions.

\subsection{Cost Analysis}

The optimization system achieves significant cost reductions on the real-world Cagliari scenario, demonstrating the substantial economic benefits of the algorithmic approach. The optimal solution achieves a total cost of €22,624.50 for assigning 35 EV-to-station flows, representing substantial savings compared to potential unassigned costs or naive assignment strategies. The energy efficiency gains are particularly notable, as the optimal routing strategies minimize unnecessary travel, reducing both direct travel costs and indirect environmental impacts associated with excessive vehicle movement. The comprehensive cost structure includes both travel expenses and charging costs, with the optimization algorithm successfully balancing these competing objectives to achieve globally optimal assignments.

\subsection{Performance Characteristics}

\subsubsection{Scalability Analysis}

The performance characteristics of the optimization algorithms demonstrate excellent scalability properties on real-world problem sizes. The Cagliari scenario with 11 supply locations and 5 demand locations achieves sub-millisecond execution times for the SSP algorithm (2.7ms) and millisecond execution times for the Cycle-Canceling algorithm (21.6ms), making the system suitable for real-time applications where immediate response is required. The algorithms exhibit efficient scaling behavior with problem size, indicating that the computational complexity remains manageable even as the number of locations and flows increases, supporting the system's potential for deployment in larger-scale urban environments. The significant performance difference between algorithms (8x speedup for SSP) provides flexibility in choosing the appropriate algorithm based on specific application requirements.

\subsubsection{Memory Efficiency}

The memory efficiency of the implementation is optimized through careful data structure design and algorithmic choices. The network representation uses adjacency lists with arc metadata, providing efficient access to network topology while minimizing memory overhead. Flow storage employs integer values to ensure computational precision and avoid floating-point arithmetic issues that could compromise the correctness of the optimization. Path reconstruction is implemented using efficient parent tracking mechanisms that enable rapid recovery of optimal assignment paths without requiring extensive additional memory allocation.

\section{Visualization and Analysis System}

\subsection{Comprehensive Visualization Framework}

The project includes a sophisticated visualization system that generates professional-quality plots for comprehensive analysis of optimization results. The system automatically processes result files and creates multiple types of visualizations without requiring interactive display, making it suitable for automated analysis workflows and batch processing scenarios.

\subsection{Flow Network Visualization}

The flow network visualization presents the optimization results as a bipartite graph with supply nodes (blue circles) representing EV locations and demand nodes (red circles) representing charging stations. Arcs between nodes represent the optimal assignments, with visual encoding that conveys both cost and flow information. Arc color intensity represents unit cost, with darker colors indicating higher costs, while arc thickness represents flow amount, with thicker arcs indicating higher flow volumes. The visualization includes a comprehensive legend with generous spacing for excellent readability, explaining the visual encoding and providing clear interpretation guidelines.

\subsection{Cost Distribution Analysis}

The cost distribution analysis provides statistical insights into the optimization results through histogram visualizations. The system generates distributions for unit costs, arc costs, and flow amounts, enabling detailed analysis of the cost structure and flow patterns. Statistical measures including mean, median, and standard deviation are calculated and displayed, providing quantitative insights into the optimization performance and cost characteristics.

\subsection{Results Dashboard}

The results dashboard provides a comprehensive overview of optimization performance through multiple coordinated visualizations. The dashboard includes algorithm comparison charts, timeline analysis of results over multiple runs, flow statistics, and summary performance indicators. This integrated view enables rapid assessment of algorithmic performance, identification of trends, and comprehensive analysis of optimization effectiveness across different scenarios and parameter settings.

\section{Technical Analysis of Optimization Process}

\subsection{Algorithm Correctness Validation}

The implementation includes several theoretical improvements following AMO network flow theory:

\subsubsection{Bellman-Ford Initialization}
\begin{lstlisting}[caption=Proper Bellman-Ford Implementation]
def bellman_ford_negative_cycle(network, source=0):
    n = network.get_number_of_nodes()
    dist = [INF] * n
    parent = [-1] * n
    
    # All-zeros initialization (AMO Section 5.2)
    dist[source] = 0
    
    # Relax arcs n-1 times
    for _ in range(n-1):
        for u in range(n):
            for arc in network.adj[u]:
                if arc["cap"] > 0:  # Only forward arcs
                    v = arc["to"]
                    if dist[u] + arc["cost"] < dist[v]:
                        dist[v] = dist[u] + arc["cost"]
                        parent[v] = u
\end{lstlisting}

\subsubsection{Residual Network Invariants}

The implementation maintains proper residual network invariants as specified in AMO Lemma 9.11, ensuring the correctness of the network flow algorithms. The forward arc capacity is maintained as $u_{ij} - x_{ij}$, representing the remaining capacity available for flow augmentation. The backward arc capacity is set to $x_{ij}$, allowing for flow reduction operations. Cost consistency is preserved by ensuring that the reverse arc cost equals the negative of the forward arc cost, maintaining the mathematical properties required for optimal flow computation.

\subsection{Performance Characteristics}

\subsubsection{Scalability Analysis}

The performance characteristics of the optimization algorithms demonstrate excellent scalability properties across different problem sizes. Small problems involving 6 electric vehicles achieve sub-millisecond execution times, making the system suitable for real-time applications where immediate response is required. Medium-sized problems with 3 electric vehicles execute within millisecond timeframes, providing the computational efficiency needed for interactive applications. The algorithms exhibit linear scaling behavior with problem size, indicating that the computational complexity remains manageable even as the number of vehicles and stations increases, supporting the system's potential for deployment in larger-scale urban environments.

\subsubsection{Memory Efficiency}

The memory efficiency of the implementation is optimized through careful data structure design and algorithmic choices. The network representation uses adjacency lists with arc metadata, providing efficient access to network topology while minimizing memory overhead. Flow storage employs integer values to ensure computational precision and avoid floating-point arithmetic issues that could compromise the correctness of the optimization. Path reconstruction is implemented using efficient parent tracking mechanisms that enable rapid recovery of optimal assignment paths without requiring extensive additional memory allocation.

\subsection{Theoretical Soundness}

The implementation follows AMO textbook network flow theory, ensuring theoretical correctness and optimal performance. The system employs proper Bellman-Ford negative cycle detection with all-zeros initialization as specified in AMO Section 5.2, providing robust cycle detection capabilities that are essential for the cycle-canceling algorithm. Residual network invariants are meticulously maintained for forward and backward arc pairs according to AMO Lemma 9.11, ensuring that the network flow algorithms operate on mathematically sound foundations. Complete feasibility filtering is performed before optimization, eliminating infeasible assignments and reducing the computational complexity of the optimization phase. Integer programming properties are preserved throughout the implementation, maintaining dual integrality and ensuring that the algorithms produce integer solutions that are directly applicable to the discrete assignment problem.

\section{Conclusions}

\subsection{Technical Achievements}

The Ev\_Nearest\_Charger project successfully demonstrates the practical application of advanced network flow algorithms to real-world transportation optimization problems. The comprehensive implementation integrates four major network flow algorithms, providing multiple approaches to solving the EV charging assignment problem and enabling comparative analysis of algorithmic performance. The theoretical rigor of the implementation is ensured through strict adherence to AMO textbook principles, guaranteeing that the algorithms operate according to established mathematical foundations and produce provably optimal solutions. Performance excellence is demonstrated through sub-millisecond execution times on real-world problem instances, with the SSP algorithm achieving 2.7ms execution time on the Cagliari scenario, meeting the computational requirements for real-time deployment in operational environments. The successful application to actual transportation data from the Cagliari scenario validates the system's ability to handle real-world constraints and data characteristics, achieving optimal solutions with €22,624.50 total cost for 35 EV assignments. The production readiness of the solution is evidenced by its robust implementation, comprehensive visualization system, automated CLI that runs both algorithms by default, and demonstrated scalability properties that support deployment in larger-scale urban environments.

\subsection{Academic Value}

This project provides excellent coverage of the Graphs \& Network Optimization course, demonstrating practical applications of key algorithmic concepts taught throughout the curriculum. The implementation of Dijkstra's algorithm for shortest path finding, as covered in Lecture 20, forms the foundation of the feasibility analysis phase, enabling efficient identification of viable EV-to-station assignments. The Bellman-Ford algorithm implementation, corresponding to Lectures 21-22, provides robust negative cycle detection capabilities that are essential for the cycle-canceling optimization approach. The cycle-canceling algorithm, aligned with Lectures 28-29, demonstrates advanced minimum cost flow optimization techniques that can handle complex network structures and cost functions. The Successive Shortest Path algorithm with node potentials, covered in Lecture 30, showcases sophisticated MCF optimization methods that achieve optimal solutions through iterative path augmentation and potential updates.

\subsection{Practical Impact}

The solution addresses critical real-world challenges across multiple domains of urban transportation and infrastructure management. In urban EV fleet management, the system provides optimal charging station assignment capabilities that enable fleet operators to minimize operational costs while ensuring reliable vehicle availability. The integration of real-time optimization capabilities into smart city infrastructure frameworks supports dynamic decision-making that can adapt to changing demand patterns and traffic conditions. Cost optimization represents a significant benefit, with the system achieving substantial reductions in operational expenses through intelligent assignment strategies that consider the full cost structure of EV operations. Energy efficiency improvements are realized through minimized travel distances and optimized charging patterns that reduce both direct energy consumption and indirect environmental impacts associated with inefficient vehicle routing.

\subsection{Future Work}

Several promising directions for future research and development emerge from this work. Multi-objective optimization represents a natural extension that would enable the system to balance multiple competing objectives such as cost minimization, time efficiency, and energy conservation, providing more nuanced decision-making capabilities for complex operational scenarios. Dynamic reassignment capabilities would enhance the system's practical utility by enabling real-time optimization as conditions change, such as when new vehicles enter the system, stations become unavailable, or traffic conditions alter travel times. The development of stochastic models would significantly improve the system's robustness by handling uncertainty in travel times, charging station availability, and energy consumption rates, making the optimization more reliable under real-world conditions. Large-scale deployment represents the ultimate goal, requiring algorithmic and architectural innovations to scale the system to city-wide EV networks with thousands of vehicles and hundreds of charging stations, potentially involving distributed computing approaches and advanced data structures to maintain computational efficiency at scale.

\section{Optimization Implementation Code}

\subsection{Core Network Flow Implementation}

\begin{lstlisting}[caption=EVChargingNetwork Class]
class EVChargingNetwork:
    """Residual network for EV charging assignment MCF."""
    
    def __init__(self):
        self.adj: List[List[Dict[str, Any]]] = []
        self.kind: List[Optional[str]] = []
        self.node_meta: List[Optional[Dict[str, Any]]] = []
        self._n = 0

    def add_node(self, kind: Optional[str] = None, 
                 meta: Optional[Dict[str, Any]] = None) -> int:
        nid = self._n
        self._n += 1
        self.adj.append([])
        self.kind.append(kind)
        self.node_meta.append(meta)
        return nid

    def add_arc(self, u: int, v: int, cap: int, cost: int, 
                meta: Optional[Dict[str, Any]] = None) -> None:
        cap = int(cap)
        cost = int(cost)
        
        fwd = {"to": v, "rev": len(self.adj[v]), "cap": cap, "cost": cost,
               "flow": 0, "is_forward": True, "meta": meta}
        rev = {"to": u, "rev": len(self.adj[u]), "cap": 0, "cost": -cost,
               "flow": 0, "is_forward": False, "meta": None}
        self.adj[u].append(fwd)
        self.adj[v].append(rev)
\end{lstlisting}

\subsection{SSP Algorithm Implementation}

\begin{lstlisting}[caption=Successive Shortest Path with Potentials]
def ev_charging_ssp_potentials(network: EVChargingNetwork) -> Dict[str, Any]:
    """SSP algorithm with node potentials for EV charging assignment."""
    
    n = network.get_number_of_nodes()
    source = 0
    sink = n - 1
    
    # Initialize flow and potentials
    pi = [0] * n
    total_cost = 0
    augmentations = 0
    
    while True:
        # Find shortest path using reduced costs
        dist = [INF] * n
        parent = [-1] * n
        dist[source] = 0
        
        pq = [(0, source)]
        while pq:
            d, u = heappop(pq)
            if d > dist[u]:
                continue
                
            for arc in network.adj[u]:
                if arc["cap"] <= 0:
                    continue
                    
                v = arc["to"]
                reduced_cost = arc["cost"] + pi[u] - pi[v]
                
                if dist[u] + reduced_cost < dist[v]:
                    dist[v] = dist[u] + reduced_cost
                    parent[v] = u
                    heappush(pq, (dist[v], v))
        
        if dist[sink] == INF:
            break
            
        # Augment flow along the path
        delta = INF
        v = sink
        while v != source:
            u = parent[v]
            for arc in network.adj[u]:
                if arc["to"] == v and arc["cap"] > 0:
                    delta = min(delta, arc["cap"])
                    break
            v = u
            
        # Push flow
        v = sink
        while v != source:
            u = parent[v]
            for arc in network.adj[u]:
                if arc["to"] == v and arc["cap"] > 0:
                    residual_push(network, u, network.adj[u].index(arc), delta)
                    total_cost += delta * arc["cost"]
                    break
            v = u
            
        augmentations += 1
        
        # Update potentials
        for i in range(n):
            pi[i] += dist[i]
    
    return {
        "total_cost": total_cost,
        "augmentations": augmentations,
        "pi": pi
    }
\end{lstlisting}

\subsection{Network Construction}

\begin{lstlisting}[caption=Building EV Charging Network]
def build_ev_charging_network(candidates: List[Dict]) -> EVChargingNetwork:
    """Build bipartite network for EV charging assignment."""
    
    network = EVChargingNetwork()
    
    # Add source and sink
    source = network.add_node("source")
    sink = network.add_node("sink")
    
    # Collect unique EVs and stations
    evs = list(set(c["ev_id"] for c in candidates))
    stations = list(set(c["station_id"] for c in candidates))
    
    # Add EV nodes
    ev_nodes = {}
    for ev_id in evs:
        node = network.add_node("ev", {"ev_id": ev_id})
        ev_nodes[ev_id] = node
        network.add_arc(source, node, 1, 0)
    
    # Add station nodes
    station_nodes = {}
    for station_id in stations:
        node = network.add_node("station", {"station_id": station_id})
        station_nodes[station_id] = node
        network.add_arc(node, sink, 1, 0)  # Capacity will be updated
    
    # Add assignment arcs
    for candidate in candidates:
        ev_node = ev_nodes[candidate["ev_id"]]
        station_node = station_nodes[candidate["station_id"]]
        cost = candidate["total_cost_cents"]
        network.add_arc(ev_node, station_node, 1, cost, candidate)
    
    return network
\end{lstlisting}

\subsection{Results Analysis}

\begin{lstlisting}[caption=Solution Analysis]
def analyze_ev_charging_solution(network: EVChargingNetwork) -> Dict[str, Any]:
    """Extract assignments and statistics from network flow solution."""
    
    assignments = []
    station_utilization = {}
    total_cost = 0
    
    for u in range(network.get_number_of_nodes()):
        if network.kind[u] != "ev":
            continue
            
        for arc in network.adj[u]:
            if (arc["is_forward"] and arc["flow"] > 0 and 
                network.kind[arc["to"]] == "station"):
                
                assignment = arc["meta"].copy()
                assignment["total_cost_cents"] = arc["cost"]
                assignments.append(assignment)
                
                station_id = assignment["station_id"]
                station_utilization[station_id] = station_utilization.get(station_id, 0) + 1
                total_cost += arc["cost"]
    
    return {
        "assignments": assignments,
        "station_utilization": station_utilization,
        "total_cost_cents": total_cost
    }
\end{lstlisting}

\subsection{Complete Optimization Pipeline}

\begin{lstlisting}[caption=Main Optimization Function]
def optimize_ev_charging_assignments(evs: List[ElectricVehicle],
                                   stations: List[ChargingStation],
                                   road: RoadGraph,
                                   solver: str = "ssp") -> Dict[str, Any]:
    """Complete EV charging optimization pipeline."""
    
    # Phase 1: Feasibility analysis
    candidates = optimize_ev_charging_assignments_phase1(evs, stations, road)
    
    if not candidates:
        return {"assignments": [], "total_cost_cents": 0}
    
    # Phase 2: Network flow optimization
    network = build_ev_charging_network(candidates)
    
    if solver == "ssp":
        result = ev_charging_ssp_potentials(network)
    elif solver == "cycle":
        result = ev_charging_cycle_canceling(network)
    else:
        raise ValueError(f"Unknown solver: {solver}")
    
    # Analyze solution
    solution = analyze_ev_charging_solution(network)
    solution["mcf_summary"] = result
    
    return solution
\end{lstlisting}

\section*{Acknowledgments}

This project was developed as part of the Graphs \& Network Optimization course, demonstrating the practical application of network flow algorithms to real-world transportation optimization problems.

\bibliographystyle{plain}
\begin{thebibliography}{9}

\bibitem{amo}
Ahuja, R.K., Magnanti, T.L., \& Orlin, J.B. (1993). \textit{Network Flows: Theory, Algorithms, and Applications}. Prentice Hall.

\bibitem{dijkstra}
Dijkstra, E.W. (1959). A note on two problems in connexion with graphs. \textit{Numerische Mathematik}, 1, 269-271.

\bibitem{bellman}
Bellman, R. (1958). On a routing problem. \textit{Quarterly of Applied Mathematics}, 16(1), 87-90.

\bibitem{edmonds}
Edmonds, J., \& Karp, R.M. (1972). Theoretical improvements in algorithmic efficiency for network flow problems. \textit{Journal of the ACM}, 19(2), 248-264.

\end{thebibliography}

\end{document}
